---
title: "Final Project 2016: Development Builds Resilience"
output: html_document
---
# Development Builds Resilience

# Overview and Motivation
National governments and global leaders are rallying behind a common goal to end global hunger and poverty by 2030, perhaps best embodied by the UN's 17 Sustainable Development Goals (SDGs), the most comprehensive effort yet to tackle the toughest development challenges facing the global community. Bread for the World supports the analysis of leaders like the World Bank and the World Food Program, which insists that getting extreme hunger and malnutrition to the 'zero zone' by 2030 is achievable in most countries. Yet there remains a small group of fragile states where unique vulnerabilities make ending hunger much more difficult. Those vulnerabilities can be categorized into three groups, known here as the "three Cs": conflict, corruption, and climate change. Countries vary in their level of vulnerability on each of the 3 Cs. Some may only be vulnerable in only one or two areas, while others could be vulnerable in all three. A rating index already exists for measuring country vulnerability on each of the three Cs. 

# Related Work
As students at the Harvard T.H. Chan School of Public Health, we are passionate about improving health outcomes for the global population.  Thus the idea of prodcing work that could help achieve the Millennium Development Goal to "end hunger, achieve food security and improved nutrition and promote sustainable agriculture" inspired us.  Additional conversations with Derek Schwabe, a research associate at Bread for the World Institute, added to our enthusiasm for this topic.

Additional reading we did included the following websites :

Bread For The World: http://www.bread.org/

Millennium Development Goals: http://www.un.org/millenniumgoals/

As well as the articles http://www.bread.org/blog/hunger-fragile-states-where-start and https://www.wider.unu.edu/publication/vulnerability-hunger and http://www.worldvision.org/news-stories-videos/what-are-fragile-states.

# Initial Goals
Using data on hunger and hunger predictors such as GDP in addition to data on climate, corruption, and conflict we wanted to achieve the following goals:

a) Create a dataset that combines measures on the 3Cs as well as on hunger that is in a useable format for organizations such as Bread for the World

b) Create a single metric that quantifies and summarizes the three Cs per country

c) Assess the relationship between hunger predictors, such as the Human Development Index, and the three Cs at the country- and continent-level 

d) Determine which of the 3Cs are most predictive of hunger

# Building a Dataset
Our analysis is comprised of the following datasets:

a) **HARMONIZED_LIST_OF_FRAGILE_SITUATIONS_FY15.csv** which is the 2014 World Bank list of countries that are deemed fragile states (low-income countries with particularly weak policies, institutions, and governance).  You can read more about how this list is created here: http://pubdocs.worldbank.org/pubdocs/publicdoc/2015/7/586581437416356109/FCS

b) **gain.csv** which is the ND-GAIN Country Index, which summarizes a country's vulnerability to climate change and other global challenges in combination with its readiness to improve resilience. All scores are between 0 and 100. A higher score means a country has more capacity to deal with potential climate risk. More information can be found here: http://index.gain.org/ranking 

We also include the **gain_delta.csv** dataset which is the GDP adjusted ND-GAIN scores.  Again, more details can be found here: http://index.gain.org/ranking 

c) **GlobalHungerIndexScore.csv** which are the 2015 Global Hunger Index Scores.  These scores are meant to measure and track hunger globally, regionally, and by country. The scores are measured on a 100-point scale where 0 is the best score (no hunger) and 100 the worst. You can read more about this dataset here: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/JL16EW

d) **2015_CPI_data.csv** which is the 2015 Corruption Perceptions Index.  A country or territory’s score indicates the perceived level of public sector corruption on a scale of 0 (highly corrupt) to 100 (very clean). More information can be found here: http://www.transparency.org/cpi2015#results-table 

e) **LowIncomeFoodDeficitCountries.csv** which is the list of the 2015 Low-Income Food-Deficit Countries.  An explanation of how this list is created can be found here:http://www.fao.org/countryprofiles/lifdc/en/

```{r}
library(readr)
library(ggplot2)
library(tidyr)
library(sjmisc)
library(dplyr)
library(plyr)
library(knitr)
library(stringr)
library(countrycode)
#####################################################################
############################ IMPORT DATA ############################
#####################################################################
#This is World Bank list of data on fragile - 2014
harmonized_list<-read_csv('HARMONIZED_LIST_OF_FRAGILE_SITUATIONS_FY15.csv')
#Clean rows
harmonized_list<-harmonized_list[(seq(2,to=nrow(harmonized_list),by=2)),]
harmonized_list<-harmonized_list[-nrow(harmonized_list),] 

# The ND-GAIN Country Index summarizes a country's vulnerability to climate change and 
nd_scores <- read_csv('gain.csv')
#GDP adjusted ND-GAIN scores.  
nd_adjusted_scores <- read_csv('gain_delta.csv')

#2015 Global Hunger Index Scores
global_hunger_index<-read_csv('GlobalHungerIndexScore.csv')

#CORRUPTION PERCEPTIONS INDEX 2015 (CPI)
corruption_index<-read_csv('2015_CPI_data.csv')

#Low-Income Food-Deficit Countries  2015
LIFDC<-read_csv('LowIncomeFoodDeficitCountries.csv')
```
We then subset the data to the most recent year available as we believe that this is most important and relevant data when thinking forward to achieving the 'zero zone' by 2030.
```{r}
#Subset vulnerability to climate change to 2014
nd_adj_2014 <- nd_adjusted_scores %>% select(ISO3,`2014`)
#Subset hunger scale 2015 
global_hunger_index_2015 <- global_hunger_index %>% select(Country,`2015`)
#Corruption_index is already just for 2015
corruption_index_subset<- corruption_index %>% select(CPI2015,wbcode)
```
Now we aggregate all the data into one dataframe.  We used the "countrycode" package for a list of all country names and codes.  This R package has data on 273 countries.

```{r}
#NOTE: The following will be indicators: fragile states (harmonized_list), Low-Income Food-Deficit Countries (LIFCD)
#NOTE: The following are on a scale:  vulnerability to climate change (nd_adjusted_scores, n=192), hunger scale (global_hunger_index, n=131)
# corruption perception index (corruption_index, n=168)

#Get all country names, codes, regions from package countrycode (data on 273 countries)
df <- countrycode_data %>% select(country.name,continent,region,iso3c,wb) %>% 
  dplyr::rename(Country = country.name,ISO3=iso3c,wbcode=wb)

#Decided we wanted to use non-GDP-adjusted ND gain scores so now we join them in
climate <- nd_scores %>% select(ISO3, 22)
df <- left_join(df, climate, by="ISO3") %>% dplyr::rename(Climate_NominalScore=`2014`)

#Merge in vulnerability to climate change (nd_adjusted_scores, n=192 but only 180 with 2014 score); checked and merge is perfect
df <- df %>% left_join(nd_adj_2014,by='ISO3') %>% dplyr::rename(Climate_AdjustedScore=`2014`)

#Merge in hunger scale (global_hunger_index_2015, n=131, only 117 have 2015 data) - had to modify some country names to match in the .csv file
df <- df %>% left_join(global_hunger_index_2015 ,by='Country') %>% dplyr::rename(Hunger_Score=`2015`)

#Merge in corruption perception index (corruption_index, n=168, all have 2015 data) - had to modify some wb codes to match in the .csv file
df <- df %>% left_join(corruption_index_subset ,by='wbcode') %>% dplyr::rename(Corruption_Score=CPI2015)
#We add in South Sudan as a country in this merge so we need to update the name correctly
df[229, 'Country'] = 'South Sudan'

#Merge in fragile states (harmonized_list) of 33 countries; Indicator=1 if it is a fragile state
df$Fragile_State <- ifelse(df$Country %in% harmonized_list$Country, 1, 0)

#Merge in Low-Income Food-Deficit Countries (LIFCD) of 54 countries; indicator of if LIFCD or not
df$LIFDC <- ifelse(df$Country %in% LIFDC$Country, 1, 0)
```
Based on conversations with our TA (Joe) and Derek, we then decided to add other datasets to this dataframe as well. These datasets helped create a new predictor variable, conflict, as well as provide alternative outcomes that serve as proxies to hunger (GDP and the Human Development Index). 

f) **Conflict.csv** which is the Global Conflict Risk Index (GCRI) developed by the European Union.  We created a score that was the average of four primary risk sub-factors that were ranked on 0-10 scale: international conflict, national conflict, unemployment (as proxy for population disgruntlement), and repression.  More information can be found at: http://conflictrisk.jrc.ec.europa.eu/

g) **GDP.csv** which is the gross domestic product (GDP) based on purchasing-power-parity (PPP) valuation of country.  The units are Current international dollar (in billions). The data was obtained from the IMF website.  

h) **HDI_table.csv** which is 2015 Human Development Index (HDI) data.  It is a composite statistic of life expectancy, education, and income per capita indicators.

```{r}
#Read in the conflict data and create an average conflict score
conf <- read_csv("Conflict.csv")
conf <- conf %>% filter(YEAR==2014) %>% select(ISO, YEAR, COUNTRY, CON_INT, CON_NB, UNEMP, REPRESS)
conf$Average_Conflict <- (conf$CON_INT + conf$CON_NB + conf$UNEMP + conf$REPRESS)/4
names(conf)[1] <- "ISO3"
conf_sub <- conf %>% select(ISO3, Average_Conflict)
#Join the average conflict score to the data frame
df <- left_join(df, conf_sub, by="ISO3")  #df is now 274 by 12

# Add in GDP data
# Gross domestic product based on purchasing-power-parity (PPP) valuation of country GDP  
# The units are Current international dollar (in billions)  
gdp<- read_csv('GDP.csv')
gdp_subset<-gdp%>% select(Country,`2015`)
df <- df %>% left_join(gdp_subset,by="Country") %>% dplyr::rename(GDP=`2015`)
df$GDP<-as.numeric(df$GDP)

#Add in HDI data
hdi<- read_csv('HDI_table.csv')
df <- df %>% left_join(hdi,by="Country") 
```
Then we created standardized values so that all scales were comparable since variables were measured on different scales.
```{r}
#Make z-scores (standardized and centered so everything is on the same scale)
df$Corruption_Zscore <- as.numeric(scale(df$Corruption_Score, center=TRUE, scale=TRUE))
df$Climate_Zscore <- as.numeric(scale(df$Climate_AdjustedScore, center=TRUE, scale=TRUE))
df$Conflict_Zscore <- as.numeric(scale(df$Average_Conflict, center=TRUE, scale=TRUE))
#Ensure all z-scores are in the correct format
df$Corruption_Zscore <- as.vector(df$Corruption_Zscore)
df$Climate_Zscore <- as.vector(df$Climate_Zscore)
df$Conflict_Zscore <- as.vector(df$Conflict_Zscore)
df$Hunger_Score <- as.numeric(df$Hunger_Score)
df$Hunger_Zscore <- as.numeric(scale(df$Hunger_Score, center=TRUE, scale=TRUE))
df$Climate_Nominal_Zscore <- as.numeric(scale(df$Climate_NominalScore, center=TRUE, scale=TRUE))
df$Conflict_Zscore_Reversed <- df$Conflict_Zscore*-1
```
Then we created summary scores based on the 3 Cs. We use z-scores to standardize all data on the 3 Cs so that they are comparable across different scales. Thus, the summary score is the average of these (or some combination of these) standardized values for corruption, conflict, and climate change. These summary scores can be interpreted as: higher values mean countries perform "above average" on the combination of the 3 C's relative to other countries; lower values mean countries perform "below average" relative to other countries.

1) **Summary_ZScore** uses the corruption Z-score, climate Z-score (based on the GDP adjusted ND-GAIN), and the conflict score which is based on the average conflict metric

2) **Summary_ZScore_2** the difference here is that the climate Z-score is based on the non-adjusted ND GAIN index

3) **Summary_ZScore_3** is a summary score created by removing Climate from Summary_ZScore_2

4) **Summary_ZScore_4** is a summary score created by removing Corruption from Summary_ZScore_2

5) **Summary_ZScore_5** is a summary score created by removing Conflict from Summary_ZScore_2

We end up using the second summary metric as we wanted to look at the relationship betwen GDP and the three C metric and thus don't want to have previously adjusted for GDP.

```{r}
#Make a summary statistic for the three Cs
df$Summary_ZScore <- (df$Corruption_Zscore+df$Climate_Zscore+df$Conflict_Zscore_Reversed)/3
#Summary statistic using non GDP adjusted climate score. What we end up using.
df$Summary_ZScore_2 <- (df$Corruption_Zscore+df$Climate_Nominal_Zscore+df$Conflict_Zscore_Reversed)/3

############### Toying with Summary Score ###############
#Removing Climate from Summary Score 
df$Summary_ZScore_3 <- (df$Corruption_Zscore+df$Conflict_Zscore_Reversed)/2

#Removing Corruption from Summary Score 
df$Summary_ZScore_4 <- (df$Climate_Nominal_Zscore+df$Conflict_Zscore_Reversed)/2

#Removing Conflict from Summary Score 
df$Summary_ZScore_5 <- (df$Corruption_Zscore+df$Climate_Nominal_Zscore)/2
```
So our data set, called df, has dimensions 274 by 25.  The columns are:

Country

continent             

region                

ISO3                 

wbcode               

Climate_NominalScore: based on the GAIN index (2014 data) 

Climate_AdjustedScore: based on the ND adjusted GAIN index (2014 data) 

Hunger_Score: Global Hunder Index Score (2015 data) 

Corruption_Score: Corruption perceptions index (2015 data) 

Fragile_State: indicator of if this is a fragile state (1) or not (0); this is the World Bank list (2014 data)

LIFDC: indicator of if this is a Low-Income Food-Deficit country (2015 data)

Average_Conflict: which is an average of four primary risk sub-factors of conflict: international conflict, national conflict, unemployment (as proxy for disgruntlement), and repression from the Global Conflict Risk Index (2015 data)

GDP: GDP based on purchasing-power-parity valuation of country (2015 data)  

HDI_2014: Human Development Index (2014 data)

Corruption_Zscore: standardized Corruption_Score        

Climate_Zscore: standardized Climate_AdjustedScore   
             
Conflict_Zscore: standardized Average_Conflict

Conflict_Zscore_Reversed: standardized Average_Conflict, reverse scale so that high=better (to align with worse-to-better scale of climate and corruption)

Hunger_Zscore: standardized Hunger_Score 

Climate_Nominal_Zscore: standardized Climate_NominalScore            

Summary_ZScore: uses the corruption Z-score, climate Z-score (based on the GDP adjusted ND-GAIN), and the conflict score which is based on the average conflict metric

Summary_ZScore_2: the difference here is that the climate Z-score is based on the non-adjusted ND GAIN index

Summary_ZScore_3: removing Climate from Summary_ZScore_2

Summary_ZScore_4 <- removing Corruption from Summary_ZScore_2

Summary_ZScore_5 <- removing Conflict from Summary_ZScore_2

```{r}
#Save to .RData file
save(df, file = "combinedData.RData")
```

# Exploratory Analysis
We did some exploratory data analysis by looking at the relationship of various predictors (climate, GDP, etc) with the hunger score. 
```{r}
############# Ggplot viz of Conflict against Hunger Score #######################
ggplot(na.omit(df), aes(Average_Conflict, Hunger_Score)) +
  geom_point(aes(color=continent)) +
  geom_smooth() +
  xlab ('Average conflict with higher score indicating more conflict') +
  ylab ('Global Hunger Score') +
  ggtitle ('Measuring Conflict Against the Global Hunger Index by Country')

############# Ggplot viz above subset / facet wrapped by continent #######################
ggplot(na.omit(df), aes(Average_Conflict, Hunger_Score)) +
  geom_point() +
  geom_smooth(method="lm") +
  facet_grid(. ~ continent) +
  xlab ('Average conflict with higher score indicating more conflict') +
  ylab ('Global Hunger Score') +
  ggtitle ('Measuring Conflict Against the Global Hunger Index by Continent')

############# Ggplot viz of Corruption against Hunger Score #######################
ggplot(na.omit(df), aes(Corruption_Score, Hunger_Score)) +
  geom_point(aes(color=continent)) +
  geom_smooth() + 
  xlab ('Corruption Score with higher score indicating less corruption') +
  ylab ('Global Hunger Score') +
  ggtitle ('Measuring Perceived Corruption Against the Global Hunger Index by Country')

############# Ggplot viz above subset / facet wrapped by continent #######################
ggplot(na.omit(df), aes(Corruption_Score, Hunger_Score)) +
  geom_point() +
  geom_smooth(method="lm") + 
  facet_grid(. ~ continent) +
  xlab ('Corruption Score with higher score indicating less corruption') +
  ylab ('Global Hunger Score') +
  ggtitle ('Measuring Perceived Corruption Against the Global Hunger Index by Continent')

############# Ggplot viz of Climate against Hunger Score #######################
climate_plot <- ggplot(na.omit(df), aes(Climate_NominalScore, Hunger_Score))
climate_plot + geom_point(aes(color=continent, size=Summary_ZScore_2)) + labs(title="Measuring the Climate Score (ND-GAIN Country Index) against the Global Hunger Index", x="ND-GAIN Country Index", y="Global Hunger Score") + geom_smooth() + scale_size(range = c(0, 3))

############# Ggplot viz of above facet wrapped by continent #######################
climate_plot + geom_point(aes(size=Summary_ZScore_2)) + geom_smooth(method="lm") + facet_grid(~continent) + labs(title="Measuring the Climate Score against the Global Hunger Index by Continent", x="ND-GAIN Country Index", y="Global Hunger Score") + scale_size(range = c(0, 3))

############# Ggplot viz of GDP against Hunger Score #######################
gdp_plot <- ggplot(na.omit(df), aes(GDP, Hunger_Score))
gdp_plot + geom_point(aes(color=continent, size=Summary_ZScore_2)) + labs(title="Measuring the GDP based on PPP against the Global Hunger Index", x="GDP based on purchasing-power-parity (PPP) valuation (in billion of current international dollars)", y="Global Hunger Score") + geom_smooth() + scale_size(range = c(0, 3))

############# Ggplot viz of GDP against Hunger Score (zoom in) #########################
df_2 <- df %>% filter(GDP<1000)
gdp_plot <- ggplot(na.omit(df_2), aes(GDP, Hunger_Score))
gdp_plot + geom_point(aes(color=continent, size=Summary_ZScore_2)) + labs(title="Measuring the GDP based on PPP against the Global Hunger Index", x="GDP based on purchasing-power-parity (PPP) valuation (in billion of current international dollars)", y="Global Hunger Score") + geom_smooth() + scale_size(range = c(0, 3))


############# Ggplot viz of above facet wrapped by continent #######################
gdp_plot + geom_point(aes(size=Summary_ZScore_2)) + geom_smooth(method="lm") + facet_grid(~continent) + labs(title="Measuring the GDP based on PPP against the Global Hunger Index by Continent", x="GDP based on purchasing-power-parity (PPP) valuation (in billion of current international dollars)", y="Global Hunger Score") + scale_size(range = c(0, 3))
```

Conflict Against the Global Hunger: Interestingly the relationship changes by continent.  For Africa and Asia, as average conflict increases, the global hunger index increases (meaning more hunger).  But for the Americas and Europe, the inverse relationship holds; as average conflict increases, the hunger decreases.

Perceived Corruption Against the Global Hunger: As corruption decreases, we see that hunger decreases.  We see this in both the overall plot and the plot broken down by continent.

Climate Score Against Global Hunger: As the ND-GAIN index increases (increase in capacity to deal with potential climate risk), the global hunger decreases. We see this in both the overall plot and the plot broken down by continent.

GDP Against Global Hunger: We can see that the trend is hard to detect given the outliers such as China and India with GDPs that are so much larger than the majority of other countries.  We also zoomed in on GDPs < 1000 billion of current international dollars. It is easier to see now that for countries with very low GDPs, there is a relationship between hunger decreasing as GDP increases.  This is more easily seen when we look by continent.  The trend is less clear in Asia, due to the outlying countries with very high GDPs.


# Exploring the Relationship Between the 3 Cs and Hunger Score
Initally we were interested in looking at the relationship between these 3Cs and the hunger score. As a reminder, the hunger score is the 2015 Global Hunger Index where 0 is the best score (no hunger) and 100 the worst. We preformed both graphical and regression analyses to investigate this association.
```{r}
###################################################
### Plotting summary score against Hunger Score####
###################################################

# Visualizations of 3C's against Hunger Score
hunger_summary <- ggplot(na.omit(df), aes(Summary_ZScore_2, Hunger_Score))
hunger_summary + geom_point(aes(color=continent)) + geom_smooth() + labs(title="Measuring the '3C' Summary Score against the Global Hunger Index by Country", x="Climate, Corruption, and Conflict Summary Score (standardized)", y="Global Hunger Score") + xlim(-1.0, 0.5)

# Subset visualizations by continent
hunger_summary + geom_point() + geom_smooth(method="lm") + facet_grid(~continent) + labs(title="Measuring the '3C' Summary Score against the Global Hunger Index by Country", x="Climate, Corruption, and Conflict Summary Score (standardized)", y="Global Hunger Score") + xlim(-1.0, 0.5)

#####################################################
### Regressing summary score against Hunger Score ###
#####################################################

# Simple regression
mod1 <-lm(Hunger_Score~Summary_ZScore_2,data=na.omit(df))
summary(mod1)

# Regression subset by continent
library(plyr)
# Break up df by continent, then fit the specified model to each piece and return a list
models <- dlply(na.omit(df), "continent", function(df) 
  lm(Hunger_Score~Summary_ZScore_2, data = na.omit(df)))
# Apply coef to each model and return a data frame
ldply(models, coef)
# Print the summary of each model
l_ply(models, summary, .print = TRUE)

###################################################
####### Regress each C against Hunger Score #######
###################################################

# Simple regression
mod2<-lm(Hunger_Score~Corruption_Score+Climate_NominalScore+Average_Conflict, data=na.omit(df))
summary(mod2)



# Break up df by continent, then fit the specified model to each piece and return a list
models2 <- dlply(na.omit(df), "continent", function(df) 
  lm(Hunger_Score~Corruption_Score+Climate_NominalScore+Average_Conflict, data = na.omit(df)))
# Apply coef to each model and return a data frame
ldply(models2, coef)
# Print the summary of each model
l_ply(models2, summary, .print = TRUE)

############### Toying with Summary Score ###############
#Summary Score is no longer a significant predictor of Hunger Score, once you remove Climate
fit<-lm(Hunger_Score~Summary_ZScore_3,data=na.omit(df))  
summary(fit)

#Summary Score is still a significant predictor of Hunger Score, even when Corruption is removed
fit<-lm(Hunger_Score~Summary_ZScore_4,data=na.omit(df))  
summary(fit)

#Summary Score is still a significant predictor of Hunger Score, even when Conflict is removed
fit<-lm(Hunger_Score~Summary_ZScore_5,data=na.omit(df))  
summary(fit)
```

When we visalize this relationship we see that it appears that as the summary score increases, the hunger score decreases. This trend is true across continents.  When we regress Hunger Score against the three 3Cs we find the coefficient of the 3Cs is statistically significant at the 0.05 alpha level.  There is a negative association between these variables

When we regress Hunger Score against the three 3Cs on a per-continent basis we find that the coefficient is significant for all continents except for Europe.  We believe that this is due to the small number of countries for which we have data on this continent.  For all continents there is a negative association between these variables.

When we do a regression of corruption, climate, and conflict as separate predictors for Hunger Score, we find that only climate is statistically significant.  There is a negative association between Hunger Score and climate.  When we regress Hunger Score against the three 3Cs separately on a per-continent basis we find that the coefficient for climate is significant for all continents (except Europe) whereas the others (Corruption and Conflict) are not.  The coefficient for climate is negative for all continents.  We see that climate is the most important predictor of Hunger Score.

When using Hunger Score as the outcome, we see that when the summary score only consists of Conflict and Corruption, it no longer significantly predicts Hunger Score.  But when you have Climate and Corruption or Climate and Conflict, the summary score is still a significant predictor of Hunger Score. We believe this suggests that Climate is the main driver in our Summary score.

# Exploring the Relationship Between the 3 Cs and HDI
When speaking with Derek, he suggested that the Human Development Index might also be a good indicator of hunger as well.  Thus we investigated the association between the 3 Cs (both cummulatively and separately) with HDI.

```{r}
############# Ggplot viz of 3C (summary score) against HDI #######################
hdi_summary <- ggplot(na.omit(df), aes(HDI_2014, Summary_ZScore_2))
hdi_summary + geom_point(aes(color=continent)) + 
  labs(title="Measuring HDI in 2014 against the '3 C' Summary Score by Country", x="Human Development Index (HDI) in 2014", y="Climate, Corruption, and Conflict Summary Score (standardized)") + geom_smooth()

############# Ggplot viz above subset / facet wrapped by continent #######################
hdi_summary + labs(title="Measuring HDI in 2014 against the '3 C' Summary Score by Continent", x="Human Development Index (HDI) in 2014", y="Climate, Corruption, and Conflict Summary Score (standardized)") + 
  geom_smooth(method="lm") + facet_grid(~continent) + geom_point(aes(size=Summary_ZScore_2)) + scale_size(range = c(0, 3))

############################ Regress 3C against HDI ######################################
fit<-lm(HDI_2014~Summary_ZScore_2,data=na.omit(df)) 
summary(fit)

############### Subset regress 3C against outcome - by continent ###############
library(plyr)
# Break up df by continent, then fit the specified model to each piece and return a list
models <- dlply(na.omit(df), "continent", function(df) 
  lm(HDI_2014~Summary_ZScore_2, data = na.omit(df)))
# Apply coef to each model and return a data frame
ldply(models, coef)
# Print the summary of each model
l_ply(models, summary, .print = TRUE)

############### Simple regression of 3 variables separately against HDI ###############
fit<-lm(HDI_2014~Corruption_Score+Climate_NominalScore+Average_Conflict, data=na.omit(df))  
summary(fit)
############### Subset regress of 3 vars against outcome - by continent ###############
# Break up df by continent, then fit the specified model to each piece and return a list
models <- dlply(na.omit(df), "continent", function(df) 
  lm(HDI_2014~Corruption_Score+Climate_NominalScore+Average_Conflict, data = na.omit(df)))
# Apply coef to each model and return a data frame
ldply(models, coef)
# Print the summary of each model
l_ply(models, summary, .print = TRUE)

############### Toying with Summary Score ###############
#Summary Score is still a significant predictor of HDI, once you remove Climate
fit<-lm(HDI_2014~Summary_ZScore_3,data=na.omit(df))  
summary(fit)

#Summary Score is still a significant predictor of HDI, even when Corruption is removed
fit<-lm(HDI_2014~Summary_ZScore_4,data=na.omit(df))  
summary(fit)

#Summary Score is still a significant predictor of HDI, even when Conflict is removed
fit<-lm(HDI_2014~Summary_ZScore_5,data=na.omit(df))  
summary(fit)
```
When we regress HDI against the three 3Cs we find the coefficient of the 3Cs is statistically significant at the 0.05 alpha level.  There is a positive association between these variables.

When we regress HDI against the three 3Cs on a per-continent basis we find that the coefficient is significant for all continents except for Europe.  We believe that this is due to the small number of countries for which we have data on this continent.  For all continents there is a positive association between these variables.

When we do a regression of corruption, climate, and conflict as separate predictors for HDI, we find that both corruption and climate are statistically significant.  There is a positive association between HDI and climate and a negative association between HDI and corruption.

When we regress HDI against the three 3Cs separately on a per-continent basis we find that the coefficient for climate is significant for all continents whereas the others are not (apart from corruption in the Americas).  Finding that climate is the most important predictor of HDI is kinda surprising but also kinda cool! 

When the summary score only consists of Conflict and Corruption, it still significantly predicts HDI. Similarly, when you consider Climate and Corruption or Climate and Conflict and HDI, the summary score is still a significant predictor. Although the effects estimates are strongest when Climate is still included in the summary score, these results also suggest that conflict and corruption may have a synergistic effect on HDI.

# Exploring the Relationship Between the 3 Cs and GDP
When speaking with Derek, he suggested that the GDP might also be a good indicator of hunger as well.  Thus we investigated the association between the 3 Cs (both cummulatively and separately) with GDP.

There are several ways to measure GDP. We used the GDP measure based on purchasing power parity (PPP) which allows one to make more accurate comparisons of the economies of two countries. In short, it is calculated by determining what each item purchased in a country would cost if it were sold in the international currency and then these are added up for all the final goods and services produced in that country for that given year. 

```{r}
############# Ggplot viz of 3C (summary score) against GDP ##########
# Plot GDP against summary score
gdp_summary <- ggplot(na.omit(df), aes(GDP, Summary_ZScore_2))
gdp_summary + geom_point(aes(color=continent)) + 
  labs(title="Measuring GDP based on PPP against the '3 C' Summary Score by Country", x="GDP based on purchasing-power-parity (PPP) valuation (in billion of current international dollars)", y="Climate, Corruption, and Conflict Summary Score (standardized)") + geom_smooth()

############# Ggplot viz above subset / facet wrapped by continent ##########
gdp_summary + labs(title="Measuring GDP based on PPP against the '3 C' Summary Score by Continent", x="GDP based on purchasing-power-parity (PPP) valuation (in billion of current international dollars)", y="Climate, Corruption, and Conflict Summary Score (standardized)") + 
  geom_smooth(method="lm") + facet_grid(~continent) + geom_point(aes(size=Summary_ZScore_2)) + scale_size(range = c(0, 3))

############################ Regress 3C against GDP #################
fit<-lm(GDP~Summary_ZScore_2,data=na.omit(df))
summary(fit)

############### Subset regress 3C against outcome - by continent ###############
library(plyr)
# Break up df by continent, then fit the specified model to each piece and return a list
models <- dlply(na.omit(df), "continent", function(df) 
  lm(GDP~Summary_ZScore_2, data = na.omit(df)))
# Apply coef to each model and return a data frame
ldply(models, coef)
# Print the summary of each model
l_ply(models, summary, .print = TRUE)

############### Simple regression of 3 variables separately against GDP ##########
fit<-lm(GDP~Corruption_Score+Climate_NominalScore+Average_Conflict, data=na.omit(df))  
summary(fit)

############### Subset regress of 3 vars against outcome - by continent ###############
# Break up df by continent, then fit the specified model to each piece and return a list
models <- dlply(na.omit(df), "continent", function(df) 
  lm(GDP~Corruption_Score+Climate_NominalScore+Average_Conflict, data = na.omit(df)))
# Apply coef to each model and return a data frame
ldply(models, coef)
# Print the summary of each model
l_ply(models, summary, .print = TRUE)
```
When we regress GDP against the three 3Cs we find the coefficient of the 3Cs is not statistically significant at the 0.05 alpha level, suggesting that the 3Cs summary measure is not predictive of GDP. 

When we do a regression of corruption, climate, and conflict as separate predictors for GDP, we find that both conflict and climate are statistically significant.  There is a positive association between GDP and climate as well as with conflict. Given that the p-value of corruption is very large when we do a regression of corruption and GDP, this is likely the driver behind why our summary score was not predictive of GDP. 

When we regress GDP against the three 3Cs separately on a per-continent basis we find the following: For Africa both conflict and climate have a statistically significant positive association with GDP. For the Americas conflict has a statistically significant positive association with GDP. For Asia and Europe we find that none of the Cs has a statistically significant association with GDP.


Looking at the plots we note several outliers in GDP which we believe may be affecting our results.  Thus we subset the data to only look at those countries with GDP of 5000 or less and re-run the regressions. This only removes three countries from the regressions: China, India, USA. <is this next statement ok?> We believe these outliers are okay to remove as they are not fragile states or countries that need focus on for the goal of ending global hunger and poverty by 2030. <again, not sure this is ok to say? india?>
```{r}
df_2<-df %>% filter(GDP<=5000)
############################ Regress 3C against GDP #################
fit<-lm(GDP~Summary_ZScore_2,data=na.omit(df_2))
summary(fit)

############### Subset regress 3C against outcome - by continent ###############
library(plyr)
# Break up df by continent, then fit the specified model to each piece and return a list
models <- dlply(na.omit(df_2), "continent", function(df) 
  lm(GDP~Summary_ZScore_2, data = na.omit(df_2)))
# Apply coef to each model and return a data frame
ldply(models, coef)
# Print the summary of each model
l_ply(models, summary, .print = TRUE)

############### Simple regression of 3 variables separately against GDP ##########
fit<-lm(GDP~Corruption_Score+Climate_NominalScore+Average_Conflict, data=na.omit(df_2))  
summary(fit)

############### Subset regress of 3 vars against outcome - by continent ###############
# Break up df by continent, then fit the specified model to each piece and return a list
models <- dlply(na.omit(df_2), "continent", function(df) 
  lm(GDP~Corruption_Score+Climate_NominalScore+Average_Conflict, data = na.omit(df_2)))
# Apply coef to each model and return a data frame
ldply(models, coef)
# Print the summary of each model
l_ply(models, summary, .print = TRUE)
```
Again, when we regress GDP against the three 3Cs we find the coefficient of the 3Cs is statistically significant at the 0.05 alpha level. And when we regress GDP against the three 3Cs on a per-continent basis we find that the 3C coefficient is significant for all continents.

Again, when we do a regression of corruption, climate, and conflict as separate predictors for GDP, we find that both conflict and climate are statistically significant.  Also when we regress GDP against the three 3Cs separately on a per-continent basis we find that conflict and climate have a statistically significant positive association with GDP for all countries.

Again we investigated playing around with the summary score.  We notice that no combination of two measures aids in the prediction of GDP. Even when corruption is removed and the summary score only includes conflict and climate, which we found to be indepenent predictors of GDP, statistical significance is still not reached. 

```{r}
############### Toying with Summary Score ###############
#Summary Score is not a significant predictor of GDP, once you remove Climate
fit<-lm(GDP~Summary_ZScore_3,data=na.omit(df))  
summary(fit)

#Summary Score is not a significant predictor of GDP, even when Corruption is removed
fit<-lm(GDP~Summary_ZScore_4,data=na.omit(df))  
summary(fit)

#Summary Score is not a significant predictor of GDP when Conflict is removed
fit<-lm(GDP~Summary_ZScore_5,data=na.omit(df))  
summary(fit)
```

#Final Analysis

Our analysis combines three unique measurements of issues facing the world today - climate change, corruption, and conflict - in order to explain several outcomes related to global hunger: a global hunger score, the Human Development Index, and GDP. What we find is that when we combine the three measurements into a single metric, the trend is similar across the board: better performance in all "three Cs" is associated with lower global hunger by country - as assessed directly by the hunger score and by proxy through HDI. When we dissect that score into its original components, we see slight variations: climate change is the strongest predictor of the hunger score and the Human Development Index, while conflict and climate are both strong predictors of GDP, but conflict is ultimately the strongest. Countries with resilience to climate change are, on average, more resilient against global hunger and are better performing developmentally. Despite the individual pieces, we have learned that, on the whole, the "big picture" summary score may provide a simple and accessible way of measuring country fragility, given that clear trends are seen using this metric. Alternative outcome measures such as the "poverty headcount ratio at $1.90/day" should also be explored, as we were unable to detect statistically significant associations between the 3Cs summary score and GDP. Nonetheless, we believe that the creation of summary scores like this may aid policymakers in setting agendas for action against global hunger, and in creating priority interventions targeted at countries who perform poorly on these three Cs. 

